### 2016-11-01 submission.tex
created outline with hipster ipsum

%!TEX program = xelatex

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Diversity, Segregation, and Complexity in Cities}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b,1]{Philip Chodrow}

\affil[a]{MIT Human Mobility and Networks Laboratory}
\affil[b]{MIT Operations Research Center}

% Please give the surname of the lead author for the running footer
\leadauthor{Chodrow}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
% \authorcontributions{Please provide details of author contributions here.}
\authordeclaration{The author declares no competing interest.}
% \equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: pchodrow@gmail.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Information Theory $|$ Diversity and Segregation $|$ Urban Theory $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{I}nformation theory...blah blah blah.

\section*{Results}

Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas. Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party. Bicycle rights kinfolk man braid disrupt XOXO. Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch. Celiac 3 wolf moon kitsch tote bag bespoke pop-up. Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb. Paleo you probably haven't heard of them whatever typewriter normcore.

Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica. Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral. Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic. Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick. Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage. Retro next level organic, photo booth letterpress typewriter deep v prism. Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.

Fanny pack poutine trust fund cardigan. Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr. Pickled gastropub vegan etsy. Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi. Roof party fixie pok pok, cardigan prism next level woke lomo. Venmo butcher irony, single-origin coffee gastropub cliche tumblr. Fap pour-over banjo, freegan literally direct trade etsy.

Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch. Kombucha whatever pug microdosing pitchfork. Activated charcoal ethical affogato echo park. Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine. Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole. Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr. Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.

\subsection*{Using some damn math} 
There's lots of damn math we could use.

\section*{Discussion}

Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas. Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party. Bicycle rights kinfolk man braid disrupt XOXO. Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch. Celiac 3 wolf moon kitsch tote bag bespoke pop-up. Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb. Paleo you probably haven't heard of them whatever typewriter normcore.

Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica. Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral. Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic. Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick. Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage. Retro next level organic, photo booth letterpress typewriter deep v prism. Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.

Fanny pack poutine trust fund cardigan. Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr. Pickled gastropub vegan etsy. Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi. Roof party fixie pok pok, cardigan prism next level woke lomo. Venmo butcher irony, single-origin coffee gastropub cliche tumblr. Fap pour-over banjo, freegan literally direct trade etsy.

Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch. Kombucha whatever pug microdosing pitchfork. Activated charcoal ethical affogato echo park. Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine. Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole. Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr. Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.

Narwhal blog poke, paleo post-ironic vegan fam butcher bushwick gastropub cardigan tofu listicle migas. Hoodie literally asymmetrical fingerstache, stumptown 90's gochujang tbh gentrify meh ethical four dollar toast plaid roof party. Bicycle rights kinfolk man braid disrupt XOXO. Leggings blue bottle post-ironic enamel pin tacos, hexagon brunch. Celiac 3 wolf moon kitsch tote bag bespoke pop-up. Master cleanse cliche try-hard bushwick everyday carry, williamsburg farm-to-table slow-carb. Paleo you probably haven't heard of them whatever typewriter normcore.

Readymade 8-bit blog church-key, fixie chicharrones succulents vape helvetica. Blue bottle aesthetic taxidermy, shabby chic la croix lumbersexual woke sartorial keytar humblebrag viral. Kickstarter humblebrag +1 tote bag, williamsburg cred 3 wolf moon pop-up glossier iceland etsy shabby chic. Vexillologist PBR B tousled humblebrag irony, kombucha pickled craft beer post-ironic synth bushwick. Godard tumeric glossier dreamcatcher humblebrag, art party tbh copper mug pok pok organic tattooed raclette sartorial ramps selvage. Retro next level organic, photo booth letterpress typewriter deep v prism. Keffiyeh pork belly hot chicken fashion axe, tumblr biodiesel truffaut bushwick edison bulb food truck iceland gochujang dreamcatcher waistcoat la croix.

Fanny pack poutine trust fund cardigan. Narwhal microdosing vape bespoke health goth humblebrag green juice, lomo chambray air plant yr. Pickled gastropub vegan etsy. Next level messenger bag meditation, health goth tacos prism snackwave subway tile hell of blog slow-carb put a bird on it lomo kogi. Roof party fixie pok pok, cardigan prism next level woke lomo. Venmo butcher irony, single-origin coffee gastropub cliche tumblr. Fap pour-over banjo, freegan literally direct trade etsy.

Church-key chicharrones pitchfork distillery, selvage kogi neutra lumbersexual lyft vexillologist waistcoat shoreditch. Kombucha whatever pug microdosing pitchfork. Activated charcoal ethical affogato echo park. Tacos live-edge sartorial, VHS brunch occupy cold-pressed blog af air plant poutine. Cardigan tumeric gochujang, drinking vinegar hot chicken scenester before they sold out tilde heirloom cornhole. Gluten-free lyft pickled irony, swag listicle subway tile jianbing schlitz disrupt distillery tumblr. Stumptown actually chartreuse authentic subway tile, pop-up pug readymade.

% \subsubsection*{Different table}

% Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

% \subsubsection*{Different table}

% Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

% \subsubsection*{SI Datasets}

% Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

% \subsubsection*{SI Movies}

% Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

% \subsubsection{Still images}

% Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.

% \subsubsection*{Appendices}

% PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{P.C. is grateful to Marta C. Gonz\'{a}lez for helpful discussion, and to MIT-Saudi Arabia...}

\showacknow{} % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% Uncomment \pnasbreak to view the references in the PNAS-style
% If you see unexpected formatting errors, try commenting out \pnasbreak
% as it can run into problems with floats and footnotes on the final page.
% \pnasbreak

% Bibliography
\bibliography{pnas-sample}

\end{document}