### 2017-08-23 8_time.rmd
cleanup

---
title: "8_time"
author: "phil"
date: "7/24/2017"
output: html_document
---

Overall goal: visualization in which we see boundaries shifting in time, and in which the importance of each boundary also changes in a way that we mark visually.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1234)
```

```{r}

library(compx)        # main analysis functions
library(tidyverse)    # for data manipulation and viz
library(igraph)       # for network analysis
library(RColorBrewer) # for plots
library(ggthemes)     # for theme_map
library(units)        # needed for sf
library(sf)           # needed for io and visualization in this vignette
library(scales)       # for ggplot2 viz
library(RSpectra)     # fast eigenvalues, also used within compx
library(tidygraph)
source('R/utils.R')
```

```{r}

tracts_to_use <- detroit_race %>% 
	group_by(tract, t) %>%
	summarise(n = sum(n)) %>%
	filter(n > 100)

tracts <- detroit_tracts %>% 
	filter(GEOID %in% tracts_to_use$tract)

```

```{r}
js_dist <- function(n,m){

	N <- sum(n)
	M <- sum(m)

	p <- n/N
	q <- m/M

	sqrt((N/(M+N))*DKL(p, p+q) + (M/(M+N))*DKL(q, p+q))
}

```

# Changing Boundaries

```{r}

race_temporal <- detroit_race %>%
	filter(t %in% c(1990, 2000, 2010))

g <- construct_information_graph(tracts, race_temporal, divergence = js_dist)

```

```{r}
k     <- 7
sigma <- 100
g     <- g %>% spectral_cluster(sigma = sigma, k = k, nreps = 1000)
```

```{r}

nodes    <- data_frame(geoid = V(g)$geoid, t = V(g)$t, cluster = V(g)$cluster)

pal <- scales::brewer_pal(palette = 'Set2')
pal <- colorRampPalette(pal(k))
pal <- pal(k)

tracts %>% 
	left_join(nodes, by = c('GEOID' = 'geoid')) %>% 
	ggplot() + 
	geom_sf(aes(fill = as.character(cluster)), size = .1) +
	ggthemes::theme_map() + 
	scale_fill_manual(values = pal) + 
	theme(legend.position = 'bottom',
		  legend.justification = c(.5, .0),
		  plot.margin=unit(c(0,0,0,0),"mm")) + 
	guides(fill = guide_legend(title = 'Cluster', nrow = 1)) + 
	facet_wrap(~t)
```

```{r}
clusters <- data_frame(GEOID = V(g)$geoid, t = V(g)$t, cluster = V(g)$cluster)

blocked_tracts <- tracts %>% 
    left_join(clusters) %>% 
    select(cluster, t) %>% 
    aggregate(by = list(cluster = .$cluster, t = .$t), FUN = identity) %>% 
    mutate(GEOID = as.character(cluster))

blocked_data <- race_temporal %>% 
    left_join(clusters, by = c('tract' = 'GEOID', 't' = 't')) %>% 
    group_by(cluster, group, t) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    mutate(tract = as.character(cluster))
```

```{r}
h <- construct_information_graph(blocked_tracts, blocked_data, js_dist)
a <- h_clust(h)
```

```{r}
get_MI <- function(k, the_year){
    v <- a %>% cutree(k)

    cluster_df <- data_frame(small_cluster = names(v), h_cluster = as.character(v)) %>% 
        tidyr::separate(small_cluster, into = c('cluster', 't'), sep = '_')

    cluster_lookup <- data_frame(name = V(g)$name, cluster = V(g)$cluster) %>% 
        separate(name, into = c('geoid', 't'), sep = '_') %>% 
        mutate(cluster = as.character(cluster)) %>% 
        left_join(cluster_df, by = c('t' = 't', 'cluster' = 'cluster')) %>% 
        mutate(t = as.integer(t))

    clustered_data <- race_temporal %>% 
        left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't'))

    clustered_data %>% 
        filter(t == the_year) %>% 
        select(-tract, cluster) %>% 
        group_by(group, h_cluster) %>% 
        summarise(n = sum(n)) %>% 
        filter(!is.na(h_cluster)) %>% 
        spread(key = h_cluster, value = n) %>% 
        ungroup() %>% 
        select(-group) %>% 
        mutual_info()
}
```

```{r}
MIs <- expand.grid(t = c(1990, 2000, 2010), k = 2:7) %>% 
    dplyr::as_data_frame() %>% 
    mutate(MI = map2_dbl(k, t, get_MI))

vals_1 <- MIs %>% 
    group_by(t) %>% 
    mutate(val = ifelse(k == 2, MI, MI - lag(MI))) %>% 
    ungroup() %>% 
    group_by(k) %>% 
    mutate(sign = val - lag(val)) %>% 
    ungroup()
```

```{r}

v <- a %>% cutree(k)

    cluster_df <- data_frame(small_cluster = names(v), h_cluster = as.character(v)) %>% 
        tidyr::separate(small_cluster, into = c('cluster', 't'), sep = '_')

    cluster_lookup <- data_frame(name = V(g)$name, cluster = V(g)$cluster) %>% 
        separate(name, into = c('geoid', 't'), sep = '_') %>% 
        mutate(cluster = as.character(cluster)) %>% 
        left_join(cluster_df, by = c('t' = 't', 'cluster' = 'cluster')) %>% 
        mutate(t = as.integer(t))

    clustered_data <- race_temporal %>% 
        left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't'))

```

```{r}

boundaries_map <- function(the_year){

all <- tracts %>% 
        aggregate(by = list(rep(1, nrow(.))), FUN = identity)

all_bd <- st_boundary(all)

m <- detroit_race %>% 
    filter(t == the_year) %>% 
    prominence_map(tracts, .)

val <- vals_1 %>% 
    filter(t == the_year)

for(j in rev(2:k)){

    v <- a %>% cutree(j)

    cluster_df <- data_frame(small_cluster = names(v), h_cluster = as.character(v)) %>% 
        tidyr::separate(small_cluster, into = c('cluster', 't'), sep = '_')

    cluster_lookup <- data_frame(name = V(g)$name, cluster = V(g)$cluster) %>% 
        separate(name, into = c('geoid', 't'), sep = '_') %>% 
        mutate(cluster = as.character(cluster)) %>% 
        left_join(cluster_df, by = c('t' = 't', 'cluster' = 'cluster')) %>% 
        mutate(t = as.integer(t))

    clustered_data <- race_temporal %>% 
        left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't'))

    plot_tracts <- tracts %>% 
        left_join(cluster_lookup, by = c('GEOID' = 'geoid')) %>% 
        aggregate(by = list(cluster = .$h_cluster, t = .$t), FUN = identity) %>% 
        filter(t == the_year)

        bd <- st_boundary(plot_tracts) %>% st_difference(all_bd)

        size <- val %>%
            filter(k == j) %>% 
            select(val) %>% 
            unlist()

        sign <- val %>% 
            filter(k == j) %>% 
            select(sign) %>% 
            unlist()

        if(the_year == 1990){
            sign = 1
        }
        if(sign > 0){
            color = 'black'
        }else{
            color = 'white'
        }

        m <- m + 
            geom_sf(size = exp(4 * size), data = bd, alpha = 0, color = color) 
}
m
}

```

```{r}
boundaries_map(1990) 
ggsave('figs/detroit_boundaries_1990.pdf', width = 4, height = 3)
```

```{r}
boundaries_map(2000) + guides(fill = FALSE)
ggsave('figs/detroit_boundaries_2000.pdf', width = 4, height = 3)
```

```{r}
boundaries_map(2010) + guides(fill = FALSE)
ggsave('figs/detroit_boundaries_2010.pdf', width = 4, height = 3)
```

```{r}
vals_1 %>% 
    ggplot() + 
    aes(x = t, y = val, color = as.factor(k), group = k) + 
    geom_point() + 
    geom_line()

```

```{r}
vals_1 %>%
    ggplot() + 
    aes(x = k, y = val, group = t, color = as.factor(t)) + 
    geom_point() + 
    geom_line() 
```

```{r}
I_CY <- vals_1 %>% 
    group_by(t) %>% 
    summarise(I_CY = sum(val))
```

```{r}
# Mutual information values
detroit_race %>% 
    spread(key = group, value = n) %>% 
    select(-tract) %>% 
    nest(-t) %>% 
    mutate(I_XY = map_dbl(data, mutual_info)) %>% 
    left_join(I_CY)
```

```{r}
detroit_race %>% 
    left_join(cluster_lookup, by = c('tract' = 'geoid', 't' = 't')) %>% 
    filter(!is.na(cluster)) %>% 
    group_by(t, group, cluster) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    group_by(cluster,t) %>% 
    mutate(perc = n / sum(n)) %>% 
    ungroup() %>%
    mutate(cluster = chartr("123456789", "ABCDEFGHI", cluster)) %>% 
    ggplot() + 
    aes(x = t, y = perc, color = group) + 
    geom_line() + 
    geom_point() + 
    facet_wrap(~cluster, nrow = 1) + 
    theme_bw() + 
    scale_color_brewer(palette = 'Set1') + 
    guides(color = FALSE) + 
    scale_x_continuous(breaks = c(1990, 2000, 2010)) + 
    theme(axis.text.x = element_text(angle = 90),
          axis.title.x=element_blank()) + 
    scale_y_continuous(labels=percent) + 
    ylab('Population Share')

ggsave('figs/detroit_breakdown.pdf', width = 10, height = 2)
```